\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual Multi-Loop Integrands

### 2023-09-22 notes/2loopNLSM.tex
small edits

Below we review the four manifestly color-dual theories that we mention in
the text, in particular self-dual Yang-Mills (SDYM) \cite{Monteiro2011pc}, two formulations of NLSM \cite{Cheung:2016prv,Cheung:2020djz,Cheung:2021zvb}, and ZM \cite{Zakharov:1973pp, Cheung:2022mix}.  All
of these theories are closely tied together by the semi-abelian YM
theory in \cref{eq:semiYM}, which gives the cubic sector of pure YM.
The manifestly color-dual equations of motion for SDYM and NLSM are
derived by picking a gauge and then placing a constraint directly on the
field strength. In almost all of these
theories, Bose symmetry is broken or obscured at the Lagrangian level.
Because of the broken Bose symmetry (and the inherited kinetic mixing
of semi-abelian YM), these theories require additional interactions to produce color-dual representations beyond one-loop.

\subsection{Self-Dual Yang-Mills} The first example in the literature
of color-dual Feynman rules is that of Yang-Mills theory in the
self-dual sector (SDYM).  We summarize the appoach of
Ref.~\cite{Monteiro2011pc} for extracting the kinematic algebra.
% \begin{equation}
% \mathcal{L}^{\text{YM}} = -\frac{1}{4}F_{\mu\nu}F^{\mu\nu}
% \end{equation}
To obtain the Lagrangian for self-dual Yang-Mills from standard Yang-Mills, we apply the
self-dual condition that equates the non-abelian $SU(N)$ field
strength,
$F_{\mu\nu} = \partial_\mu A_\nu -\partial_\nu A_\mu + g[A_\mu
,A_\nu]$, to its dual two-form:
\begin{equation}
  \label{SDcon} F_{\mu\nu} = \frac{1}{2}
  \varepsilon_{\mu\nu\rho\sigma}F^{\rho\sigma} \,.
\end{equation}
Solutions to the self-dual condition automatically satisfy the
Yang-Mills equation of motion due to the Bianchi identity:
\begin{equation}
D^\mu F^{\mu\nu} = \frac{1}{2} \varepsilon_{\mu\nu\rho\sigma}D^\mu F^{\rho\sigma} = 0 \,.
\end{equation}
Furthermore, field configurations that satisfy \cref{SDcon} correspond to instantons which are probed by the path integral in the full non-perturbative description of Yang-Mills. To construct a perturbative action, begin by
transforming from Cartesian coordinates to light cone coordinates,
$(t,x,y,z) \rightarrow (u,v,w,\bar{w})$, via:
\begin{equation}
u = t+x,\quad v=t-x, \quad w = y+iz, \quad \bar{w} = y-iz
\end{equation}
which yields the invariant line element:
\begin{equation}
ds^2 = 2(dudv - dwd\bar{w}) = dt^2 - d\vec{x}\cdot d\vec{x} \,.
\end{equation}
Applying this coordinate transformation to the self-dual condition of
\cref{SDcon}, produces three simple constraint equations:
\begin{align} \label{SDLightCone1}
F_{uw} &=0\,,
\\
 \label{SDLightCone2}
 F_{uv} &= F_{w\bar{w}}\,,
\\
 \label{SDLightCone3}
 F_{v\bar{w}} &= 0\,.
\end{align}
Thus, in light-cone gauge, $A_u=0$, we realize the self-dual
conditions in terms of the gauge field as
\begin{align}
  A_u=0 \text{ and } (\ref{SDLightCone1}) \quad&\Rightarrow \quad A_w=0\\
  A_w=0\,,\,A_u=0 \text{ and } (\ref{SDLightCone2}) \quad&\Rightarrow \quad \partial_u A_v = \partial_w A_{\bar{w}} \label{eq:sd-divs}\,.
\end{align}
This last constraint is satisfied by writing the gauge field in
terms of a single degree of freedom which we call $\Psi$:
\begin{equation}
  \label{SDDef}
A_v = \frac{1}{2} \partial_w \Psi, \qquad A_{\bar{w}} = \frac{1}{2} \partial_u \Psi \,.
\end{equation}
Using this definition on the first self-dual constraint equation
yields the following equation of motion for the dynamical scalar
field:
\begin{equation}
  \label{SDEOM}
(\ref{SDLightCone1}) \text{ and } (\ref{SDDef}) \quad  \Rightarrow\quad \Box \Psi + ig [\partial_u \Psi, \partial_w \Psi ]=0 \,.
\end{equation}
Adding back in the anti-holomorphic field $\bar{\Psi}$ as a Lagrange
multiplier, we obtain the following interaction Lagrangian for
self-dual Yang-Mills theory:
\begin{equation}
  \mathcal{L}^{\text{SDYM}} = (\partial \bar{\Psi})(\partial \Psi) -i g \bar{\Psi} [\partial_u \Psi, \partial_w \Psi ] \,.
\end{equation}
The color-ordered ordered cubic Feynman rule for this theory can be
immediately read off as:
\begin{equation}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =X(k_1,k_2) \equiv k_{u,1}k_{w,2} - k_{w,1}k_{u,2} \,.
\end{equation}
In this form, the Lagrangian and Feynman rules are manifestly
color-dual \textit{off-shell}. We can see the effect \cref{SDcon} had
on the Feynman rules of the full Yang-Mills Lagrangian: it simply
decoupled the anti-MHV three-point vertex from the theory, leaving
only the MHV three-point vertex. Before proceeding, it is important to
note that the mass-dimension of the theory appears to differ from that
of Yang-Mills -- we comment on this in detail in \cref{Conventions}.

>>>>

Below we review the four manifestly color-dual theories that we mention in
the text, self-dual Yang-Mills (SDYM) \cite{Monteiro2011pc}, two formulations of NLSM \cite{Cheung:2016prv,Cheung:2020djz,Cheung:2021zvb}, and ZM \cite{Zakharov:1973pp, Cheung:2022mix}.  All
of these theories are closely tied together by the semi-abelian YM
theory in \cref{eq:semiYM}, which gives the cubic sector of pure YM.
The manifestly color-dual equations of motion for SDYM and NLSM are
derived by picking a gauge and then placing a constraint directly on the
field strength. In almost all of these
theories, Bose symmetry is broken or obscured at the Lagrangian level.
Since Bose symmetry is broken (and inherit the kinetic mixing
of semi-abelian YM), these theories require additional interactions to produce color-dual representations beyond one-loop.

\subsection{Self-Dual Yang-Mills} The first example in the literature
of color-dual Feynman rules is that of Yang-Mills theory in the
self-dual sector (SDYM).  We summarize the appoach of
Ref.~\cite{Monteiro2011pc} for extracting the kinematic algebra.
% \begin{equation}
% \mathcal{L}^{\text{YM}} = -\frac{1}{4}F_{\mu\nu}F^{\mu\nu}
% \end{equation}
To obtain the Lagrangian for self-dual Yang-Mills from standard Yang-Mills, we apply the
self-dual condition that equates the non-abelian $SU(N)$ field
strength,
$F_{\mu\nu} = \partial_\mu A_\nu -\partial_\nu A_\mu + g[A_\mu
,A_\nu]$, to its dual two-form:
\begin{equation}
  \label{SDcon} F_{\mu\nu} = \frac{1}{2}
  \varepsilon_{\mu\nu\rho\sigma}F^{\rho\sigma} \,.
\end{equation}
Solutions to the self-dual condition automatically satisfy the
Yang-Mills equation of motion due to the Bianchi identity:
\begin{equation}
D^\mu F^{\mu\nu} = \frac{1}{2} \varepsilon_{\mu\nu\rho\sigma}D^\mu F^{\rho\sigma} = 0 \,.
\end{equation}
Furthermore, field configurations that satisfy \cref{SDcon} correspond to instantons which are probed by the path integral in the full non-perturbative description of Yang-Mills. To construct a manifestly perturbative action, one begins by
transforming from Cartesian coordinates to light cone coordinates,
$(t,x,y,z) \rightarrow (u,v,w,\bar{w})$, via:
\begin{equation}
u = t+x,\quad v=t-x, \quad w = y+iz, \quad \bar{w} = y-iz
\end{equation}
which yields the invariant line element:
\begin{equation}
ds^2 = 2(dudv - dwd\bar{w}) = dt^2 - d\vec{x}\cdot d\vec{x} \,.
\end{equation}
Applying this coordinate transformation to the self-dual condition of
\cref{SDcon}, produces three simple constraint equations:
\begin{align} \label{SDLightCone1}
F_{uw} &=0\,,
\\
 \label{SDLightCone2}
 F_{uv} &= F_{w\bar{w}}\,,
\\
 \label{SDLightCone3}
 F_{v\bar{w}} &= 0\,.
\end{align}
Thus, in light-cone gauge, $A_u=0$, we realize the self-dual
conditions in terms of the gauge field as
\begin{align}
  A_u=0 \text{ and } (\ref{SDLightCone1}) \quad&\Rightarrow \quad A_w=0\\
  A_w=0\,,\,A_u=0 \text{ and } (\ref{SDLightCone2}) \quad&\Rightarrow \quad \partial_u A_v = \partial_w A_{\bar{w}} \label{eq:sd-divs}\,.
\end{align}
This last constraint is satisfied by writing the gauge field in
terms of a single degree of freedom which we call $\Psi$:
\begin{equation}
  \label{SDDef}
A_v = \frac{1}{2} \partial_w \Psi, \qquad A_{\bar{w}} = \frac{1}{2} \partial_u \Psi \,.
\end{equation}
Using this definition on the first self-dual constraint equation
yields the following equation of motion for the dynamical scalar
field:
\begin{equation}
  \label{SDEOM}
(\ref{SDLightCone1}) \text{ and } (\ref{SDDef}) \quad  \Rightarrow\quad \Box \Psi + ig [\partial_u \Psi, \partial_w \Psi ]=0 \,.
\end{equation}
Adding back in the anti-holomorphic field $\bar{\Psi}$ as a Lagrange
multiplier, we obtain the following interaction Lagrangian for
self-dual Yang-Mills theory:
\begin{equation}
  \mathcal{L}^{\text{SDYM}} = (\partial \bar{\Psi})(\partial \Psi) -i g \bar{\Psi} [\partial_u \Psi, \partial_w \Psi ] \,.
\end{equation}
The color-ordered ordered cubic Feynman rule for this theory can be
immediately read off as:
\begin{equation}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =X(k_1,k_2) \equiv k_{u,1}k_{w,2} - k_{w,1}k_{u,2} \,.
\end{equation}
In this form, the Lagrangian and Feynman rules are manifestly
color-dual \textit{off-shell}. We can see the effect \cref{SDcon} had
on the Feynman rules of the full Yang-Mills Lagrangian: it simply
decoupled the anti-MHV three-point vertex from the theory, leaving
only the MHV three-point vertex. Before proceeding, it is important to
note that the mass-dimension of the theory appears to differ from that
of Yang-Mills -- we comment on this in detail in \cref{Conventions}.
